Planetary perturbers: Flaring star-planet interactions in Kepler and TESS

### 2023-10-23 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets \textcolor{red}{could} become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares on the stellar surface. \textcolor{red}{If and how their energy and morphology differs from intrinsic stellar flares is poorly constrained.} However, as we expect planet-induced flares to be associated with the planet's orbital period $P_{\rm orb}$, we can identify their presence from a clustering of flares in phase with the planet's orbital phase. We used the excellent phase coverage from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase. In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected power of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, consistent with low expected power. Those we expect to show planet-induced flaring fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected power of interaction increases. The flares in HIP 67522 are prominently clustered ($p<0.006$). This young Hot Jupiter system is the most promising candidate for magnetic star-planet interaction in our sample.

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits of only a few stellar radii~\citep[e.g., ][]{sanchis-ojeda2014study} that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Specifically, the effects on the star pose challenges, but also opportunities for our understanding of star-planet systems.

@@

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic SPI can occur if the planet is orbiting within the Alfv\'en radius of the star at least part of the time~\citep{preusse2006magnetic, cohen2011dynamics}. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints from large samples for calibration, which we aim to provide in this work.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been cataloged~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Since 2018, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. The key advantage of both missions for this work is the nearly uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

@@

The absolute power of flaring SPI is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates for dissipated energy~\citep{lanza2018closeby} are in the ballpark of the regime of superflares\textcolor{red}{, with energies released at optical wavelengths $\geq10^{33}\,$erg~\citep{schaefer2000superflares},} observed by Kepler and TESS. However, on the one hand, the planet might be triggering flares prematurely, which would be of lower energy~\citep{loyd2023flares}, and end up below our detection threshold. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would \textcolor{red}{unfold faster or slower than} intrinsic flares. We will assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should still have the same physics as intrinsic flares\textcolor{red}{. They should then approximalety follow the fast-rise-exponential-decay morphology. Individual flares can deviate from this template, yet still be classified as flares because flares remain the best explanation for a sudden rise and drop in stellar brightness in most main sequence dwarf stars. Many of these deviations are unexplained today~\citep{howard2022no}, so that identifying SPI flares based on morphology is not possible without additional constraints. However, if we can statistically identify enough SPI flares, as we attempt in this work, we can infer the differences between intrinsic and SPI events from the two ensembles.}

@@

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
In contrast to global SPI, local SPI effects are easier to tell apart from intrinsic stellar behavior because they are tied to the planet's orbital period. \textcolor{red}{Searches for star-planet interactions in polarized radio emission, i.e. at the wavelengths at which Jupiter-Io interaction is observed (scaled up to the high field strengths of stars) have produced promising, but tentative results. In some works, shorter radio wavelenths can still be confused with stellar activity~\citep{pineda2023coherent}. In others, the planet has not been detected (yet)~\citep{vedantham2020coherent}, the data do not cover multiple orbital periods to see the signal repeat~\citep{perez-torres2021monitoring}, or both \citep{callingham2021population}.} Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an active Hot Jupiter host, higher X-ray energy flares seem only marginally clustered in orbital phase than lower energy ones~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emission when the Jupiter-sized companion was in periastron on the highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner \mbox{TRAPPIST-1} planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring SPI. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the clustering of in orbital phase, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). We then estimated the theoretically expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date measurements. The relevant properties of the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

@@

The Transiting Exoplanet Survey Satellite~\citep[TESS,][]{ricker2015transiting} is an all-sky mission that began operations in 2018, completed two sky scans until summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

@@

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science Python package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ (or $1.5 \sigma$ for stars like AU Mic or Proxima Cen, which are very active or have very low noise levels, or both) threshold as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked. Transits are usually too shallow to affect either light curve de-trending or flare finding. The light curve portions around deep transits were inspected manually.

In the de-trended light curves, we search for flares as series of at least three consecutive data points $3\sigma$ above the noise level using the \texttt{AltaiPony} method \texttt{find\_flares}. We estimate the noise level as the standard deviation in a rolling window of two hours, while masking deviation above $2.5\sigma$ (or $1.5\sigma$  for stars like AU Mic and Proxima Cen). To capture the exponential tail of the flare, we use the \texttt{addtail} flag in the \texttt{find\_flares} method to continue adding data points to the end of the flare as long as they are $2\sigma$ above the noise threshold.

We vet all flare candidates by eye, and exclude instrumental and physical false positives, such as Solar System Objects passing by the field of view~(see Fig.~\ref{fig:flares_and_fps} in the Appendix for an example). For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
Equivalent duration is equal to the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In \textcolor{red}{short}, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. \textcolor{red}{For this, we sample 10000 times from the expected distribution the same number of flares we observe in each system, e.g., for HIP 67522, we sample 12 flares distributed according to the blue curve in Fig.~\ref{fig:cumdist_transiting}. For each of the 10000 samples, we compute the Anderson-Darling statistic $A^2$, which yields a distribution of $A^2$ values. In a last step, we compute $A^2$ for the observed distribution, and compute the $p$-value.} We repeat this test with different phase offsets to account for remaining biases in sensitivity of the test at different phases.

@@

\begin{figure}[ht!]
    \script{paper_lx_plots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_lxlbol.png}
        \caption{
           X-ray luminosity over bolometric luminosity compared to Rossby number (\textbf{upper panel}), and average surface magnetic field strength (\textbf{lower panel}). \textcolor{red}{Blue lines denote the error bars, grey lines point to the IDs of the corresponding stars.} Both relations follow those of stars not selected for hosting close-in planets, except for GJ 3082, which is underluminous for its Rossby number, and hence also its estimated magnetic field appears too strong for its $L_X/L_{\rm bol}$.
        }
        \label{fig:lx}
    \end{centering}
\end{figure}

@@

First, we estimate the power of SPI generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet}. \textcolor{red}{We adopt their Eqn. 45 for non-linear and axisymmetric magnetic fields, because the non-linear formulation does not assume a constant force-free parameter, and because only axisymmetric fields are able to produce modulation of flaring activity in phase with the planetary orbit. We note that the axisymmetry of stellar magnetic fields varies across stars~\citep[e.g.,][]{donati2009magnetic, donati2008largescale, morin2008largescale, morin2010largescale}, and also with the stellar activity cycle of individual stars~\citep[e.g.,][]{borosaikia2016solarlike, lehmann2021identifying} In Eqn. 45,} $P_{\rm spi,sb}$ is proportional to

\begin{equation}
    P_{\rm spi,sb}=\underbrace{\frac{27\pi \lambda^2(n+1)}{16 \mu_0 n (\lambda^2 + n^2)^{1/3}}}_{\hat{C}} B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3}  v_{\rm rel} \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}},
\end{equation}

\textcolor{red}{where $\mu_0$ is the magnetic permeability of the vacuum, and $n$ and $\lambda$ are dimensionless constants, all of which we subsume under $\hat{C}$ together with other constant factors}. $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=0.25$ ($\lambda^2 = 1.01203$, which best reproduces observations of SPI candidate systems HD~179949, $\tau$ Boo, and HD~189733 in~\citet{lanza2012starplanet} (except for $\upsilon$ And). Since the planetary fields are unknown, we adopt $B_{\rm p}=1\,$G for simplicity.} We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Section~\ref{sec:methods:bfield}) as a proxy.

@@

\begin{equation}
    P_{\rm spi,sb0} =\hat{C} R_{\rm p}^2 B_*^2 v_{\rm rel}  \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}}
\end{equation}

Third, we estimate the power of SPI generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, \textcolor{red}{ combining Eqn.~8 for the power of interaction and Eqn.~11 for the magnetopause radius of the planet from\citet{kavanagh2022radio} to arrive at:}

\begin{equation}
    P_{\rm spi,aw} = \pi^{1/2} B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta
\end{equation}

\textcolor{red}{Here, we assume that the magnetopause radius extends to where the magnetic pressures of planet and stellar magnetic wind are equal, and the wind is dominated by the magnetic field in de sub-Alf\'enic regime. We approximate the stellar wind density $\rho_W$ with a proportionality to $\rho_* (R_* / a)^{2}$, where $\rho_*$ is the stellar wind density at the base of the corona. For the stellar wind magnetic field at the planetary radius $B_W$, we insert $B_* (R_* / a)^{3}$, i.e., the field strength decreasing radially outward from the surface as a dipole. This produces an upper bound on the field strength at the planet's distance because more complex fields would drop off more steeply. The last two replacements result in the total $R_*^2 a^{-2}$ factor in the above equation. Since $\rho_*$ is poorly constrained empirically~\citep{vidotto2021evolution}, we adopt $\rho_*=2\cdot10^{10}\rm{cm}^{-3}$, a value often used for the Sun~\citep{sokolov2013magnetohydrodynamic}, and for stellar simulations~\citep{alvarado-gomez2020tuning,kavanagh2021planetinduced}. The angle between the wind magnetic field vector and the velocity vector of the planet $\theta$ is also unknown, so we set $\sin\theta=1$ to obtain an upper limit. In reality, $\theta$ will vary both between systems, and over time in any individual system. Omitting $\rho_*$ assumes that all stellar wind densities are the same in our sample, which is not true in reality. The dependence on $\rho_*$ is stronger than that on $B_*$, but weaker than on every other parameter, so it may dominate the magnetic field dependence.}

@@

\newpage
\section{Flaring finding in Kepler and TESS: False positives}
\restartappendixnumbering

>>>>

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets \textcolor{red}{could} become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger \textcolor{red}{flares in the corona. Potential differences to the energies and morphologies of intrinsic flares are poorly constrained.} However, as we expect planet-induced flares to \textcolor{red}{correlate} with the planet's orbital period, we can identify \textcolor{red}{them} from a clustering of flares in phase with the planet's orbit. We used the excellent phase coverage from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase. In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected power of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, consistent with low expected power. Those we expect to show clustering fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected power of interaction increases. \textcolor{red}{The flares in HIP 67522 are prominently clustered ($p<0.006$). This young Hot Jupiter system is the most promising candidate for magnetic star-planet interaction in our sample.}

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits of only a few stellar radii~\citep[e.g., ][]{sanchis-ojeda2014study}, so that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Specifically, the effects on the star pose challenges, but also opportunities for our understanding of star-planet systems.

@@

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic SPI can occur if the planet is orbiting within the Alfv\'en radius of the star at least part of the time~\citep{preusse2006magnetic, cohen2011dynamics}. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by \textcolor{red}{a planet that crosses the magnetic field} cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints from large samples for calibration, which we aim to provide in this work.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been cataloged~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. Since 2018, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. The key advantage of both missions for this work is the nearly uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

@@

The absolute power of flaring SPI is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates for dissipated energy~\citep{lanza2018closeby} are in the ballpark of the regime of superflares\textcolor{red}{, with energies released at optical wavelengths $\geq10^{33}\,$erg~\citep{schaefer2000superflares}, that can be} observed by Kepler and TESS. However, on the one hand, the planet might be triggering flares prematurely, which would be of lower energy~\citep{loyd2023flares}, and end up below our detection threshold. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would \textcolor{red}{unfold faster or slower than} intrinsic flares. We will assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should still have the same physics as intrinsic flares\textcolor{red}{. They should then approximalety follow the fast-rise-exponential-decay morphology~(see Fig.~\ref{fig:flares_and_fps} for an example). Individual intrinsic flares can deviate from this template, yet still be classified as flares because flares remain the best explanation for a sudden rise and drop in stellar brightness in most main sequence dwarf stars. Many of these deviations are unexplained today~\citep{howard2022no}, so that identifying SPI flares based on morphology is not possible without additional constraints. However, if we can statistically identify enough SPI flares, as we attempt in this work, we can infer the differences between intrinsic and SPI events from the two ensembles.}

@@

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
In contrast to global SPI, local SPI effects are easier to tell apart from intrinsic stellar behavior because they are tied to the planet's orbital period. \textcolor{red}{Searches for an orbital modulation of polarized radio emission have produced promising, but tentative results. Close-in star-planet systems are expected to interact similarly to the Jupiter-Io system~\citep{bigg1964influence}, albeit scaled up to the high field strengths of stars~\citep{turnpenney2018exoplanetinduced}. In some detections, the planet has not been found (yet)~\citep{vedantham2020coherent}, or the data do not cover multiple orbital periods to see the signal repeat~\citep{perez-torres2021monitoring, pineda2023coherent}, or both \citep{callingham2021population}.} Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an active Hot Jupiter host, higher X-ray energy flares seem only marginally clustered in orbital phase than lower energy ones~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emission when the Jupiter-sized companion was in periastron on the highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner \mbox{TRAPPIST-1} planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring SPI. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the clustering of \textcolor{red}{flares} in orbital phase, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). We then estimated the theoretically expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date measurements. The relevant properties of the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

@@

The Transiting Exoplanet Survey Satellite~\citep[TESS,][]{ricker2015transiting} is an all-sky mission that began operations in 2018, \textcolor{red}{and has} completed two sky scans in summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

@@

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science Python package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ \textcolor{red}{threshold} (or $1.5 \sigma$ for stars like AU Mic or Proxima Cen, which are very active or have very low noise levels, or both) as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked. Transits are usually too shallow to affect either light curve de-trending or flare finding. The light curve portions around deep transits were inspected manually.

In the de-trended light curves, we search for flares as series of at least three consecutive data points $3\sigma$ above the noise level using the \texttt{AltaiPony} method \texttt{find\_flares}. We estimate the noise level as the standard deviation in a rolling window of two hours, while masking deviations above $2.5\sigma$ (or $1.5\sigma$  for stars like AU Mic and Proxima Cen). To capture the exponential tail of the flare, we use the \texttt{addtail} flag in the \texttt{find\_flares} method to continue adding data points to the end of the flare as long as they are $2\sigma$ above the noise threshold.

We vet all flare candidates by eye, and exclude instrumental and physical false positives, such as Solar System Objects passing by the field of view~(see Fig.~\ref{fig:flares_and_fps} in the Appendix for an example). For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
\textcolor{red}{In other words, equivalent duration is }the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In \textcolor{red}{short}, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per \textcolor{red}{orbital phase} bin into one expected distribution \textcolor{red}{(blue curves in Figs.~\ref{fig:cumdist_transiting} and \ref{fig:cumdist_rv})} by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. \textcolor{red}{For this, we sample 10000 times from the expected distribution the same number of flares as we observe in each system. For instance, for HIP 67522, we sample 12 flares distributed according to its blue curve in Fig.~\ref{fig:cumdist_transiting}. For each of the 10000 samples, we compute the Anderson-Darling statistic $A^2$, which yields a distribution of $A^2$ values. In a last step, we compute $A^2$ for the observed distribution, and compute the $p$-value.} We repeat this test with different phase offsets to account for remaining biases in sensitivity of the test at different phases.

@@

\begin{figure}[ht!]
    \script{paper_lx_plots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_lxlbol.png}
        \caption{
           X-ray luminosity over bolometric luminosity compared to Rossby number (\textbf{upper panel}), and average surface magnetic field strength (\textbf{lower panel}). \textcolor{red}{Blue lines denote the error bars, black lines point to the IDs of the corresponding stars.} Both relations follow those of stars not selected for hosting close-in planets, except for GJ 3082, which is underluminous for its Rossby number, and hence also its estimated magnetic field appears too strong for its $L_X/L_{\rm bol}$.
        }
        \label{fig:lx}
    \end{centering}
\end{figure}

@@

First, we estimate the power of SPI generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet}. \textcolor{red}{We adopt their Eqn.~45 for non-linear and axisymmetric magnetic fields, because the non-linear formulation does not assume a constant force-free parameter, and because only axisymmetric fields can produce modulation of flaring activity in phase with the planetary orbit. We note that the axisymmetry of stellar magnetic fields varies across stars~\citep[e.g.,][]{donati2009magnetic, donati2008largescale, morin2008largescale, morin2010largescale}, and also with the stellar activity cycle of individual stars~\citep[e.g.,][]{borosaikia2016solarlike, lehmann2021identifying}. In Eqn. 45, the power of SPI, $P_{\rm spi,sb}$, is}

\begin{equation}
\textcolor{red}{      P_{\rm spi,sb}=\underbrace{\frac{27\pi \lambda^2(n+1)}{16 \mu_0 n (\lambda^2 + n^2)^{1/3}}}_{\hat{C}} B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3}  v_{\rm rel} \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}},}
\end{equation}

\textcolor{red}{where $\mu_0$ is the magnetic permeability of the vacuum, and $n$ and $\lambda$ are dimensionless constants, all of which we subsume under $\hat{C}$, together with other constant factors}. $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=0.25$ ($\lambda^2 = 1.01203$), which best reproduces observations of SPI candidate systems HD~179949, $\tau$ Boo, and HD~189733 in~\citet{lanza2012starplanet} (except for $\upsilon$ And). Since the planetary fields are unknown, we adopt $B_{\rm p}=1\,$G for simplicity.} We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Sections~\ref{sec:methods:bfield} and~\ref{sec:discussion:as}) as a proxy.

@@

\begin{equation}
\textcolor{red}{      P_{\rm spi,sb0} =\hat{C} R_{\rm p}^2 B_*^2 v_{\rm rel}  \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}}}
\end{equation}

Third, we estimate the power of SPI generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, \textcolor{red}{ combining Eqn.~8 for the power of interaction and Eqn.~11 for the magnetopause radius of the planet from~\citet{kavanagh2022radio} to arrive at:}

\begin{equation}
\textcolor{red}{    P_{\rm spi,aw} = \frac{\pi^{1/2}}{2^{2/3}} B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta}
\end{equation}

\textcolor{red}{Here, we assume that the magnetopause radius extends to where the magnetic pressures of planet and stellar magnetic wind are equal, and that the wind is dominated by the magnetic field in the sub-Alf\'enic regime. We approximate the stellar wind density $\rho_W$ with a proportionality to $\rho_* (R_* / a)^{2}$, where $\rho_*$ is the stellar wind density at the base of the corona. For the stellar wind magnetic field at the planetary radius $B_W$, we insert $B_* (R_* / a)^{3}$, i.e., we assume that the field strength decreases radially outward from the surface as a dipole. This produces an upper bound on the field strength at the planet's distance because more complex fields would drop off more steeply. The last two replacements result in the total $R_*^2 a^{-2}$ factor in the above equation. Since $\rho_*$ is poorly constrained empirically~\citep{vidotto2021evolution}, we adopt $\rho_*=2\cdot10^{10}\rm{cm}^{-3}$, a value often used for the Sun~\citep{sokolov_magnetohydrodynamic_2013}, and for stellar simulations~\citep{alvarado-gomez2020tuning,kavanagh2021planetinduced}. The dependence on $\rho_*$ is stronger than that on $B_*$, so it may dominate the magnetic field dependence. However, it is also weaker than that on any other parameter in the scaling law. The angle between the wind magnetic field vector and the velocity vector of the planet $\theta$ is also unknown, so we set $\sin\theta=1$ to obtain an upper limit. In reality, $\theta$ will vary both between systems, and over time in any individual system.}

@@

\newpage
\section{Finding flares in Kepler and TESS: False positives}
\restartappendixnumbering
